---
title: "Kelly Whelan, HW5_R"
output: html_document
---

SW: need to use ``` vs '''

'''{r}
```{r}
#Problem 1
#Figure out how to generically extract all numeric columns from any data frame so the method you develop can be used on any data. 

#calls the data frame 'diamonds' into memory 
# SW: need to load ggplot2 library to make data set available
library(ggplot2)
data(diamonds)

#'str' does not return anything, it allows you to view the structure of the data frame of 'diamonds' in this case
str(diamonds)

#'<-' assigns a value to a name, here a new vector is assigned to 'numCol' that uses the built in 'sapply' which by default returns a vector, 
#which takes all the parts of the data frame diamonds in this case and pulls out the ones that are numeric
numCol <- diamonds[sapply(diamonds, is.numeric)]

#str() of the newly defined numCol vector shows the parts of diamonds dataframe that actually are numeric
str(numCol)


#Problem 2
#Create a data frame that contains each pair of variable names in the first column in a single string separated by #a, and their corresponding Pearson correlation coefficient in the second column.

#calling the data frame of diamonds to be used
data(diamonds)

#<- assigns numCol for each of the conditions presented
#is.numeric is a more general test of an object being interpretable as numbers
numCol <- diamonds[sapply(diamonds, is.numeric)]        

#'paste' concatenates vectors after converting to character, 'cor' computes the variance of x and the correlation of x and y 
numCol <- within(numCol, "carat-depth" <- paste(cor(carat,depth,method="pearson")))

numCol <- within(numCol, "depth-table" <- paste(cor(depth,table,method="pearson")))

numCol <- within(numCol, "table-price" <- paste(cor(table,price,method="pearson")))

numCol <- within(numCol, "price-x" <- paste(cor(price,x,method="pearson")))

numCol <- within(numCol, "x-y" <- paste(cor(x,y,method="pearson")))

numCol <- within(numCol, "y-z" <- paste(cor(y,z,method="pearson")))

numCol <- subset(numCol, select = -c(1,2,3,4,5,6,7))


str(numCol)
```

```{r}
# SW: lost points: this is not correct, you should be calculating the 
# correlation for all unique pairs of numeric variables (21 total pairs) 
# and returning a data frame with a single row per pair. see below.
get_numeric <- function(dat) {
  nc <- lapply(dat, class) %in% c("numeric","integer")
  return(dat[,nc])
}

corrfn <- function(dat) {
  comb <- combn(colnames(dat), 2)
  out <- apply(comb, MARGIN = 2, 
               FUN = function(x) {
                 return(c(paste(x[1], x[2], sep="-"), 
                          cor(dat[x[1]], dat[x[2]])))
               })
  return(data.frame(vars = out[1,], corr = out[2,]))
}

corrfn(get_numeric(diamonds))
```

```{r}
#Problem 3
#Create and label a scatter plot for every pair of numeric variables. Add a title to the plot that contains the calculated Pearson correlation coefficient of variables contained in the plot.

#'require' loads and attaches add-on packages, in this case the packages of 'grid' and 'ggplot2' in order to compute the scatter plots below
require(grid)
require(ggplot2)

#'<-' assigns a value to a name, here a new vector is assigned to 'numCol' that uses the built in 'sapply' which by default returns a vector, 
#which takes all the parts of the data frame diamonds in this case and pulls out the ones that are numeric
numCol <- diamonds[sapply(diamonds, is.numeric)]
str(numCol)

#This is a scatter plot of variable 'Carat' to variable 'Depth' for the Pearson Correlation Coefficient of 0.0282243142790482.
plot(numCol$carat,numCol$depth,main = "Pearson Coeff = 0.0282243142790482", xlab = "Carat", ylab = "Depth")

#This is a scatter plot of variable 'Depth' to variable 'Table' for the Pearson Correlation Coefficient of 0.0282243142790482.
plot(numCol$depth,numCol$table,main = "Pearson Coeff = -0.295778521526582", xlab = "Depth", ylab = "Table")

#This is a scatter plot of variable 'Table' to variable 'Price' for the Pearson Correlation Coefficient of 0.0282243142790482.
plot(numCol$table,numCol$price,main = "Pearson Coeff = 0.127133902121742", xlab = "Table", ylab = "Price")

#This is a scatter plot of variable 'Price' to variable 'x' for the Pearson Correlation Coefficient of 0.0282243142790482.
plot(numCol$price,numCol$x,main = "Pearson Coeff = 0.884435161016127", xlab = "Price", ylab = "x")

#This is a scatter plot of variable 'x' to variable 'y' for the Pearson Correlation Coefficient of 0.0282243142790482.
plot(numCol$x,numCol$y,main = "Pearson Coeff = 0.974701479670349", xlab = "x", ylab = "y")

#This is a scatter plot of variable 'y' to variable 'z' for the Pearson Correlation Coefficient of 0.0282243142790482.
plot(numCol$y,numCol$z,main = "Pearson Coeff = 0.952005716156274", xlab = "y", ylab = "z")

```

```{r, eval = F}
# SW: lost points: the plots you show are correct, but you were supposed 
# to create a scatterplot for each of the pairs of numeric variables. 
# also, specifying the plots one by one is tedious and error-prone, see
# code below.
plotnum <- function(dat) {
  dat_nc <- get_numeric(dat)
  lbl <- corrfn(dat_nc)
  comb <- combn(colnames(dat_nc), 2)
  apply(comb, MARGIN = 2,
        FUN = function(x) {
          plot(dat_nc[,x[1]], dat_nc[,x[2]], xlab = x[1], ylab = x[2],
               main = paste0(x[1],"-",x[2],": ",
                             lbl[lbl[,1] == paste(x[1],x[2],sep="-"),2]))
        })
}

plotnum(diamonds)
```
SW: same issue with ''' vs ```
'''