---
title: "Kelly Whelan, HW6_R"
output: html_document
---

SW: did you test your code? it doesn't work as written. i've modified it a bit to help isolate the problem, but it will require more extensive fixes. from what I can tell just based on reading the code (vs being able to run it), it looks like you aren't returning a single data frame and your `explore` function does not accept a correlation threshold. also, within your `explore` function, you assign new values to the function object itself, it's not at all clear how this design will work.

```{r}
#Problem 1
#Plot a pair of blue histograms with a vertical red line at the mean for every numerical variable at each bin size specified in the bin sizes input parameter.

#calls the data frame 'diamonds' and 'mtcars' into memory 
# SW: you cannot load the diamonds data set if you have not first loaded
# ggplot2, this causes all subsequent references to diamonds to fail.
require(ggplot2)
require(grid)

data(diamonds)
data(mtcars)

#binsizes <- c(5, 20, 50)
#threshold <- 0.25
#df <- data(diamonds)
#Here 'A' is being assigned and 'as.logical' attempts to coerce its argument to be of logical type. 'rep' replicates the values in x. 'c' forms a vector.

diamonds <- within(diamonds, 'A' <- as.logical(rep(c("True","False"), c(23599, 30341)))) 
#str(A) will show us that logi is made up of Trues and Falses
str(diamonds)
#'<-' assigns a value to a name, here a new vector is assigned to 'var_num', 'var_factor' and 'var_logical' that uses the built in 'sapply' which by default returns a vector, which takes all the parts of the data frame diamonds in this case and pulls out the numeric, factor and logical columns
var_num <- diamonds[sapply(diamonds, is.numeric) == TRUE]
var_factor <- diamonds[sapply(diamonds, is.factor) == TRUE]  
var_logical <- diamonds[sapply(diamonds, is.logical) == TRUE]
str(var_num)
str(var_factor)
str(var_logical)
```

```{r}
#Add loop in function for all variables, bin sizes and mean line
explore <- function(df, binsizes, corr_thresh) 
{
  for (i in 1:length(var_num))
  {
    for (bins in 1:length(binsizes))
    {
      #Calculate bin size
      num_bins <- (max(var_num[[i]])-min(var_num[[i]]))/binsizes[bins]
      
      #Calculate the mean and label
      col_mean <- mean(var_num[[i]])
      mean_Label = sprintf("%8.2f ", col_mean)
      
      #Plot regular count histogram
      explore <- ggplot(var_num, aes(x=var_num[[i]]),
                         environment = environment())
      explore <- explore +
        geom_histogram(colour = 'blue',
                       fill = 'blue',
                       binwidth=num_bins) +
        labs(x=names(var_num)[[i]]) +
        geom_vline(xintercept=col_mean, colour='red') +
        annotate("text",x=col_mean,y=0,label=mean_Label,hjust=0)
      print(explore)
      #Plot denisty histogram
      print(explore + aes(y=..density..) + labs(y="Density"))
      }
    }
#Problem #2
#Plot a gray bar graph for every categorical and binary variable.  
  for ( k in 1:length(var_logical))
    {
      explore <- ggplot(var_logical, aes(x=var_logical[[k]]), 
                        environment = environment()) 
  
explore <- explore + geom_bar(fill="gray", colour="gray") + labs(x=names(var_logical)[k])
print(explore) 
      }
#Problem 3
#Calculate the r^2 value for every pair of numerical variables.
      for(i in 1:length(var_num))
      {
      r_square <- cor(var_num[i] ,var_num[i+1])^2
      print(r_square)
      }
  }
```

```{r}
#Test it
explore(diamonds, c(5,20,50))
```

```{r}


#Problem 4
#a: A frequency table for every categorical and binary variable.
freq_a <-table(var_factor[1])
freq_b <- table(var_factor)
str(freq_a)
str(freq_b)

freq_c <-table(var_logical[1])
freq_d <- table(var_logical)
str(freq_a)
str(freq_b)

```


