---
title: "Kelly Whelan, TEST2_R"
output: html_document
---

```{r}
Following is an example of how to get a dataset from the web, load it into R and begin exploring the data.
```{r}
# UCI is a great resource for datasets. Here's an example of how to 
# retrieve the Abalone data
uciaddress <- "http://archive.ics.uci.edu/ml/machine-learning-databases/"
dataset <- "abalone/abalone.data"
getdataset <- paste(uciaddress, dataset, sep="")
abalone <- read.csv(getdataset)
str(abalone)
library(gridExtra) #this package is for a later grid plot
```

Now let's name the columns:
```{r}
colnames(abalone) <- c("Gender","Length","Diameter",
                       "Height", "Whole_wgt",
                       "Shucked_wgt", "Viscera wgt",
                       "Shell wgt", "Rings")
str(abalone)
```

Here's a method for get a random sample from a dataframe

```{r}
#Problem 1

#sample sizes: 1%, 10%, 25%, 50%, 80% with 42, 418, 1044, 2088, 3341, 4176 respectively

# Quick way to get a sample from a dataframe
randomSample = function(dataframe,n,replacement) {
  return (dataframe[sample(nrow(dataframe), n, replace=replacement),]) }

#generating sample sizes from the abalone dataframe
sample_1 <- randomSample(abalone, 42, replacement=FALSE)
sample_10 <- randomSample(abalone, 418, replacement=FALSE)
sample_25 <- randomSample(abalone, 1044, replacement=FALSE)
sample_50 <- randomSample(abalone, 2088, replacement=FALSE)
sample_80 <- randomSample(abalone, 3341, replacement=FALSE)
sample_100 <- randomSample(abalone, 4176, replacement=FALSE)

#Calculate the min of the Whole_wgt column from each sample and store in a dataframe
#minimum values for each sample size

a <- min(sample_1$Whole_wgt)
b <- min(sample_10$Whole_wgt)
c <- min(sample_25$Whole_wgt)
d <- min(sample_50$Whole_wgt)
e <- min(sample_80$Whole_wgt)
f <- min(sample_100$Whole_wgt)

#vector created to store each min value
SampleMin <- c(a, b, c, d, e, f)

#Calculate the max of the Whole_wgt column from each sample and store in a dataframe
#maximum value for each sample size
a1 <- max(sample_1$Whole_wgt)
b1 <- max(sample_10$Whole_wgt)
c1 <- max(sample_25$Whole_wgt)
d1 <- max(sample_50$Whole_wgt)
e1 <- max(sample_80$Whole_wgt)
f1 <- max(sample_100$Whole_wgt)

#vector created to store each max value
SampleMax <- c(a1, b1, c1, d1, e1, f1)

#mean values for each sample size
m1 <- mean(sample_1$Whole_wgt)
m2 <- mean(sample_10$Whole_wgt)
m3 <- mean(sample_25$Whole_wgt)
m4 <- mean(sample_50$Whole_wgt)
m5 <- mean(sample_80$Whole_wgt)
m6 <- mean(sample_100$Whole_wgt)

#vector created to store each mean value
SampleMean <- c(m1, m2, m3, m4, m5, m6)

#variance values for each sample size
v1 <- var(sample_1$Whole_wgt)
v2 <- var(sample_10$Whole_wgt)
v3 <- var(sample_25$Whole_wgt)
v4 <- var(sample_50$Whole_wgt)
v5 <- var(sample_80$Whole_wgt)
v6 <- var(sample_100$Whole_wgt)

#vector created to store each variance value
SampleVar <- c(v1, v2, v3, v4, v5, v6)

#putting all the calculated vecotrs into one dataframe
df <- data.frame(SampleMin, SampleMax, SampleMean, SampleVar)

#transpose function and we get
df <- data.frame(t(df))

str(df)

#renaming the column names to reflect the sample size
colnames(df) <- c("1% Sample", "10% Sample", "25% Sample", "50% Sample", "80% Sample", "100% Sample")

str(df)

```

```{r}
#Problem 2
#Create a grid of histograms that contains all sample sizes

require(ggplot2)
require(grid)
require(gridExtra)
samplelist<-c(42,418,1044,2089,3342,4117)

abaloneweight <- data.frame(abalone$whole.weight)

histplot <- function(data,listnum)
{
  par(mfrow=c(2,3))
  for (k in 1:ncol(data))
  {
  for (i in listnum)
  {  
    hist(data[sample(1:nrow(data),i),k], main=paste("Histogram size of", i ), xlab= i)
  }
  }
}
histplot(abaloneweight,samplelist)

```

```{r}
#Problem 3
#Implement a loop that takes 1,000 samples at each of the sample sizes in 1 above (except the 100% sample!). Calculate a mean at each iteration and store the results in a dataframe. There should be five columns and 1,000 rows. Name the columns but NOT the rows

#Samps <- rep((randomSample(abalone, 42, FALSE)), 1000)
#Samps <- data.frame(Samps)
#print(Samps$Whole_wgt[1])
#for(i in 1:length(Samps)){
#  my list <- list()
#  while (i <= 1000){
#    Mean <- mean(Samps$Whole_wgt[i])
#    myList <- Mean
#  }
#}
#print(Mean)
############################################################

sample_1f=c(rep(0,times=1000))
sample_10f=c(rep(0,times=1000))
sample_25f=c(rep(0,times=1000))
sample_50f=c(rep(0,times=1000))
sample_80f=c(rep(0,times=1000))

for (i in 1:1000){
randomSample = function(dataframe,n,replacement) {
  return (dataframe[sample(nrow(dataframe), n, replace=replacement),]) }
  sample_1 <- randomSample(abalone, 42, FALSE)
  sample_10 <- randomSample(abalone, 418, FALSE)
  sample_25 <- randomSample(abalone, 1044, FALSE)
  sample_50 <- randomSample(abalone, 2088, FALSE)
  sample_80 <- randomSample(abalone, 3341, FALSE)


sample_1mean1 = mean(sample_1[,5])
sample_1f[i] = sample_1mean1 
sample_10mean1 = mean(sample_10[,5])
sample_10f[i] = sample_10mean1 
sample_25mean1 = mean(sample_25[,5])
sample_25f[i] = sample_25mean1 
sample_50mean1 = mean(sample_50[,5])
sample_50f[i] = sample_50mean1 
sample_80mean1 = mean(sample_80[,5])
sample_80f[i] = sample_80mean1 

}

dff=data.frame(cbind(sample_1f,sample_10f,sample_25f,sample_50f,sample_80f))
str(dff)
colnames(dff) <- c("1% Sample", "10% Sample", "25% Sample", "50% Sample", "80% Sample")

```

```{r}
#Problem 4
#Create another grid of histograms that contains the five columns just created and include the 100% sample from 1. 
require(ggplot2)
require(grid)

hist1<- ggplot(dff[1], aes(x=dff[[1]]))
hist1 <- hist1 + geom_histogram(colour='blue', fill='black') + labs(x=names(dff)[[1]])
print(hist1)

hist2<- ggplot(dff[2], aes(x=dff[[2]]))
hist2 <- hist2 + geom_histogram(colour="grey", fill="black") + labs(x=names(dff)[[2]])
print(hist2)

hist3<- ggplot(dff[3], aes(x=dff[[3]]))
hist3 <- hist3 + geom_histogram(colour="magenta", fill="black") + labs(x=names(dff)[[3]])
print(hist3)

hist4<- ggplot(dff[4], aes(x=dff[[4]]))
hist4 <- hist4 + geom_histogram(colour="pink", fill="black") + labs(x=names(dff)[[4]])
print(hist4)

hist5<- ggplot(dff[5], aes(x=dff[[5]]))
hist5 <- hist5 + geom_histogram(colour="red", fill="black") + labs(x=names(dff)[[5]])
print(hist5)

hist6<- ggplot(df[6], aes(x=df[[6]]))
hist6 <- hist6 + geom_histogram(colour="yellow", fill="black") + labs(x=names(df)[[6]])
print(hist6)


grid.arrange(hist1, hist2, hist3, hist4, hist5, hist6, ncol=3)


```

```{r}
#Problem 5
#Calculate the min, max, mean and variance of each column from 3 and of the total dataset (the 100% sample). Append to the dataframe created in 1. Name the mean row ResampleMean and the others appropriately.
#Calculate the min of the Whole_wgt column from each sample and store in a dataframe
#minimum values for each sample size

diffa <- min(dff[,1])
diffb <- min(dff[,2])
diffc <- min(dff[,3])
diffd <- min(dff[,4])
diffe <- min(dff[,5])
difff <- min(sample_100[,5])

#vector created to store each min value
DiffSampleMin <- c(diffa, diffb, diffc, diffd, diffe, difff)

#Calculate the max of the Whole_wgt column from each sample and store in a dataframe
#maximum value for each sample size
diffa1 <- max(dff[,1])
diffb1 <- max(dff[,2])
diffc1 <- max(dff[,3])
diffd1 <- max(dff[,4])
diffe1 <- max(dff[,5])
difff1 <- max(sample_100[,5])

#vector created to store each max value
DiffSampleMax <- c(diffa1, diffb1, diffc1, diffd1, diffe1, difff1)

#mean values for each sample size
diffm1 <- mean(dff[,1])
diffm2 <- mean(dff[,2])
diffm3 <- mean(dff[,3])
diffm4 <- mean(dff[,4])
diffm5 <- mean(dff[,5])
diffm6 <- mean(sample_100[,5])

#vector created to store each mean value
DiffSampleMean <- c(diffm1, diffm2, diffm3, diffm4, diffm5, diffm6)

#variance values for each sample size
diffv1 <- var(dff[,1])
diffv2 <- var(dff[,2])
diffv3 <- var(dff[,3])
diffv4 <- var(dff[,4])
diffv5 <- var(dff[,5])
diffv6 <- var(sample_100[,5])

#vector created to store each variance value
DiffSampleVar <- c(diffv1, diffv2, diffv3, diffv4, diffv5, diffv6)

#putting all the calculated vecotrs into one dataframe
dffff <- data.frame(DiffSampleMin, DiffSampleMax, DiffSampleMean, DiffSampleVar)

#transpose function and we get
dffff <- data.frame(t(df))

str(df)

#renaming the column names to reflect the sample size
colnames(dffff) <- c("1% Sample", "10% Sample", "25% Sample", "50% Sample", "80% Sample")
hi <- rbind(df, dffff)

hi <- data.frame(hi)
colnames(hi) <- c("1% Sample", "10% Sample", "25% Sample", "50% Sample", "80% Sample", "100% Sample")

str(dffff)


```


```{r}
#Problem 6
#Create a single histogram of the 100% data and overlay and label lines that represent the 1%, 10%, 25%, 50%, and 80% SampleMean.

#mean values for each sample size
m1 <- mean(sample_1$Whole_wgt)
m2 <- mean(sample_10$Whole_wgt)
m3 <- mean(sample_25$Whole_wgt)
m4 <- mean(sample_50$Whole_wgt)
m5 <- mean(sample_80$Whole_wgt)
#m6 <- mean(sample_100$Whole_wgt)

Handle_Num2 <- function(Num_Frame) 
{
  for (i in 1:length(Num_Frame)){
    col_mean1 <- SampleMean [[1]]
    col_mean2 <- SampleMean [[2]]
    col_mean3 <- SampleMean [[3]]
    col_mean4 <- SampleMean [[4]]
    col_mean5 <- SampleMean [[5]]
    mean_Label1 = sprintf("10f", col_mean1)
    mean_Label2 = sprintf("10f", col_mean2)
    mean_Label3 = sprintf("10f", col_mean3)
    mean_Label4 = sprintf("10f", col_mean4)
    mean_Label5 = sprintf("10f", col_mean5)

  
    Num_Plot <- ggplot(Num_Frame, aes(x=Num_Frame[[i]]),
                     environment = environment())
    Num_Plot <- Num_Plot + geom_histogram(colour ='gray', fill ='gray') + labs(x=names(Num_Frame)[[i]]) + geom_vline(xintercept=col_mean1, colour ='black', size=1) + geom_vline(xintercept=col_mean2, colour = 'black',size=1) + geom_vline(xintercept=col_mean3, colour = 'pink',size=1) + geom_vline(xintercept=col_mean4, colour = 'pink',size=1) + geom_vline(xintercept=col_mean5, colour = 'blue',size=1) + annotate("text", x=col_mean1,y=0, label ="1%=red") + annotate("text", x=col_mean2,y=0, label ="10%=black") + annotate("text", x=col_mean3,y=0, label ="25%=pink") + annotate("text", x=col_mean4,y=0, label ="50%=pink") + annotate("text", x=col_mean5,y=0, label ="80%=blue")

  print(Num_Plot)
  }
}

Handle_Num2(abalone[5])

```

```{r}
#7
#Create a second histogram of the 100% data and overlay and label vertical lines that represent the 1%, 10%, 25%, 50%, and 80% ResampleMean.

DiffSampleMean <- c(diffm1, diffm2, diffm3, diffm4, diffm5, diffm6)

Handle_Num2 <- function(Num_Frame) 
{
  for (i in 1:length(Num_Frame)){
    col_mean1 <- DiffSampleMean [[1]]
    col_mean2 <- DiffSampleMean [[2]]
    col_mean3 <- DiffSampleMean [[3]]
    col_mean4 <- DiffSampleMean [[4]]
    col_mean5 <- DiffSampleMean [[5]]
    mean_Label1 = sprintf("10f", col_mean1)
    mean_Label2 = sprintf("10f", col_mean2)
    mean_Label3 = sprintf("10f", col_mean3)
    mean_Label4 = sprintf("10f", col_mean4)
    mean_Label5 = sprintf("10f", col_mean5)

  
    Num_Plot <- ggplot(Num_Frame, aes(x=Num_Frame[[i]]),
                     environment = environment())
    Num_Plot <- Num_Plot + geom_histogram(colour ='gray', fill ='gray') + labs(x=names(Num_Frame)[[i]]) + geom_vline(xintercept=col_mean1, colour ='black', size=1) + geom_vline(xintercept=col_mean2, colour = 'black',size=1) + geom_vline(xintercept=col_mean3, colour = 'pink',size=1) + geom_vline(xintercept=col_mean4, colour = 'pink',size=1) + geom_vline(xintercept=col_mean5, colour = 'blue',size=1) + annotate("text", x=col_mean1,y=0, label ="1%=red") + annotate("text", x=col_mean2,y=0, label ="10%=black") + annotate("text", x=col_mean3,y=0, label ="25%=pink") + annotate("text", x=col_mean4,y=0, label ="50%=pink") + annotate("text", x=col_mean5,y=0, label ="80%=blue")

  print(Num_Plot)
  }
}

Handle_Num2(abalone[5])


```

```{r}
#8
#Now execute your code on diamonds$carat and faithful$eruptions.
data(faithful)

sample_1f=c(rep(0,times=1000))
sample_10f=c(rep(0,times=1000))
sample_25f=c(rep(0,times=1000))
sample_50f=c(rep(0,times=1000))
sample_80f=c(rep(0,times=1000))

for (i in 1:1000){
randomSample = function(dataframe,n,replacement) {
  return (dataframe[sample(nrow(dataframe), n, replace=replacement),]) }
  sample_1 <- randomSample(faithful, 3, FALSE)
  sample_10 <- randomSample(faithful, 27, FALSE)
  sample_25 <- randomSample(faithful, 68, FALSE)
  sample_50 <- randomSample(faithful, 136, FALSE)
  sample_80 <- randomSample(faithful, 218, FALSE)


sample_1mean1 = mean(sample_1[,1])
sample_1f[i] = sample_1mean1 
sample_10mean1 = mean(sample_10[,1])
sample_10f[i] = sample_10mean1 
sample_25mean1 = mean(sample_25[,1])
sample_25f[i] = sample_25mean1 
sample_50mean1 = mean(sample_50[,1])
sample_50f[i] = sample_50mean1 
sample_80mean1 = mean(sample_80[,1])
sample_80f[i] = sample_80mean1 

}

dff=data.frame(cbind(sample_1f,sample_10f,sample_25f,sample_50f,sample_80f))
str(dff)
colnames(dff) <- c("1% Sample", "10% Sample", "25% Sample", "50% Sample", "80% Sample")
```

```{r}
#9
#Provide a brief explanation of what you observe in the measures and histograms as it relates to sample size and resampling.

# From this assignment I observed that resampling allows data to get closer and closer to the mean. 

```